\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-02-02 doc/publications/plos16/plos16.tex
changes to results section

The model paramters are $w_i = 1/\sqrt{2^{i-1}}$ for $i =1,\ldots 5$ and we set 
$$
z =  \sum_{i=1}^{i=5} {w_i x_i}.
$$
We let $\sigma_\epsilon^2 = Var(z)(1-\theta)/\theta $  where $\theta$ is a parameter controling the fraction of variance explained by the infomative variables and in our study we chose $\theta = 0.125$ . Then 
$y =   z + \epsilon$ where $\epsilon \sim  N(0, \sigma_\epsilon^2).$
The dichotomous response is generated by thresholding $y$ at the $0.5$ quantile.

@@

We fit the random forest model and estimate the classification accuracy by capturing the out-of-bag error (OOB) and the
variable selection performance by capturing the rank-biased-overlap (RBO) \cite{Webber2010} measure. We expect to
retrieve the 5 features in order of the weight that was assigned to them and with a higher variable importance measure
than the noise variables. RBO provides a quantiative measure of this.  We are running Apache Spark 1.6.1 on a YARN
cluster with 12 worker nodes each with 16 Intel Xeon E5-2660@2.20GHz CPU cores and 128 GB of RAM.

@@

As shown in Fig~\ref{figure:out-of-bag-prediction-error-prod.png} the default value for \mtry (number of variables
selected at each node) does not result in a good classification performance for this large feature dataset. Note, that
the plot shows the proportion of $n/p$ for \mtry\ with the default being $\sqrt{p} \approx 1581$ ($n/p=0.0006$).  The
OOB for this value of \mtry does not drop below 0.5, even when the number of trees are increased (\ntree).

Increasing \mtry\ in combination with \ntree\ yields best performance with the OOB error essentially constant around 0.4
across a large range of of \mtry and \ntree values.  This is in contrast to the feature-selection performance, where the
RBO measure heavily depends on \ntree\ and gives better results with lower values of \mtry, where the best RBO is around
0.85 on a scale from 0 (no feature recovered) to 1 (fully recovered).

This may be because a large \mtry\ generates more correlated trees as the same important features have a higher chance
of being selected in all, not yielding good performance outcomes.  This issue is less pronounced for classification
tasks where random features can mimic the response variable, hence resulting more in a performance plateau.  Increasing
the number of trees on the other hand improves performance especially when the trees are kept diverse (small \mtry) but
appropriate for large feature datasets (\mtry larger than default). \cursedforest's feature of parallelising trees at
the node level hence cater perfectly for the requirement of large feature datasets as more trees can be built given a
fixed time budget.

>>>>

The model paramters are $w_i = 1/\sqrt{2^{i-1}}$ for $i =1,\ldots 5$ and we set 
\begin{equation}
\label{equation:synthetic.data}
z =  \sum_{i=1}^{i=5} {w_i x_i}.
\end{equation}
We let $\sigma_\epsilon^2 = Var(z)(1-\theta)/\theta $  where $\theta$ is a parameter controling the fraction of variance explained by the infomative variables and in our study we chose $\theta = 0.125$ . Then 
$y =   z + \epsilon$ where $\epsilon \sim  N(0, \sigma_\epsilon^2).$
The dichotomous response is generated by thresholding $y$ at the $0.5$ quantile.

@@

We fit the random forest model and estimate the classification accuracy by capturing the out-of-bag error (OOB) and the
variable selection performance by capturing the rank-biased-overlap (RBO) \cite{Webber2010} measure. RBO compares two
lists (in this case, the paramters $w_i, i=1,\ldots , 5$ (equation \ref{equation:synthetic.data}), and the top 1000
ranked variable importances from the \cursedforest model). It is a mesaure of similarity between rankings that handles
list of different length, weighs high ranks more heavily than low, and is monotonic with increasing depth of
evaluation. We expect to retrieve the 5 features in order of the weight that was assigned to them and with a higher
variable importance measure than the noise variables. RBO provides a quantiative measure of how close we come to this on
a scale from 0 (no feature recovered) to 1 (fully recovered). See supplementary information (S*) for more details.

We are running Apache Spark 1.6.1 on a YARN
cluster with 12 worker nodes each with 16 Intel Xeon E5-2660@2.20GHz CPU cores and 128 GB of RAM.

@@

As shown in Fig~\ref{figure:out-of-bag-prediction-error-prod.png} the default value for \mtry (number of variables
selected at each node) does not result in a good classification performance for this large feature dataset. Note that
the plot shows the proportion, $n/p,$ for \mtry\ with the default being $\sqrt{p} \approx 1581$ ($n/p=0.0006$).  The
OOB for this value of \mtry does not drop below 0.5, even when the number of trees are increased (\ntree).

Increasing \mtry\ in combination with \ntree\ yields best performance with the OOB error essentially constant around 0.4
across a large range of of \mtry and \ntree values.  This is in contrast to the feature-selection performance, where the
RBO measure heavily depends on \ntree\ and gives better results with lower values of \mtry.

This may be because a large \mtry\ generates more correlated trees as the same important features have a higher chance
of being selected in all, not yielding good performance outcomes.  This issue is less pronounced for classification
tasks where random features can mimic the response variable, hence resulting more in a performance plateau.  
\marginpar{unclear what this means}
Increasing
the number of trees on the other hand improves performance especially when the trees are kept diverse (small \mtry) but
appropriate for large feature datasets (\mtry larger than default). \cursedforest's feature of parallelising trees at
the node level hence cater perfectly for the requirement of large feature datasets as more trees can be built given a
fixed time budget.
